---
title: "Introducing incidence2"
description: |
  Compute, handle and visualise epicurves.
author:
  - name: Tim Taylor
date: 07-08-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
creative_commons: CC BY
bibliography: ref.bib
citation: true
---

It's been a few months since our first post, ['Welcome to the reconverse'](posts/2021-03-09-welcome-to-the-reconverse/), and, coinciding with the release of version 1.2.0, we thought it a good time to introduce you to the ***incidence2*** package [@Rincidence2] for **R** [@R].

## Background
During a disease outbreak it is important to collect some basic information on cases in a line list;
a data structure where each entry describes a unique individual in an epidemiological context.
Line lists are typically in a tabular format where each row is an individual with a unique ID,
meta-data describing the individual which are fixed in time (e.g., age, sex, occupation, disease outcome)
and time-stamped measurements (e.g., date of onset, date of admission, date of disease outcome).

To better understand an outbreak, line lists are often aggregated to produce epidemic curves (epicurves);
graphs representing the number of new cases over a given time interval (e.g. days, weeks, months).
The process to convert a line list to an epicurve typically involves four stages:

1. Data cleaning;
2. time standardisation;
3. aggregation; and
4. visualisation.

To assist with the computing and visualisation of epicurves (stages 2 to 4), the **incidence** [@Kamvar2019] package was created.
Here the term *incidence* was used to refer to the underlying, aggregated data, not the *rate of incidence*.
Whilst this original implementation worked well it did have some limitations:

* Users were unable to group by more than one variable at a time.
* Users were also unable to convert pre aggregated data in to the packages' incidence class.
* Date grouping was only possible using the functions provided within the package.
  
***incidence2*** is a re-imagining of the original package, built to address these limitations.


## Implementation

With ***incidence2***, incidence objects are constructed as a subclass of [tibble](https://tibble.tidyverse.org/) [@Rtibble], where the following *class invariants*^[structural information that must remain true for an object to maintain its class when operated on] hold true:

* A single column representing the *time index* of the latter counts;
* one or more columns representing *counts* of a time-stamped measurement;
* zero or more columns representing *groups*;
* zero or more columns representing other *associated measurements*; and
* when considering the time index and group columns, there must be no duplicate rows.

The names of the columns representing the time index, counts, groups and associated measurements are then stored as as attributes of the object.

By building incidence objects on tibbles (which themselves build upon data frames) we benefit from the rich support for these data structures that is present within both 'base' R and the wider package ecosystem.

When constructing incidence objects input can be given as either line list or pre-aggregated data and users have the option to use an "opinionated" approach to date grouping (built on top of the ***grates*** package [@Rgrates]) or to specify there own date grouping function.


## How to use
To illustrate the main package functionality we give three simple examples.
The first of these uses synthetic line list data from the ***outbreaks*** package [@Routbreaks].
The second example downloads some Covid-19 case data from the UK's National Health Service via the ***covidregionaldata*** package [@Rcovidregionaldata].
The third example uses the ***clock*** [@Rclock] and ***zoo*** [@Rzoo] packages to illustrate the use of custom date grouping functions.

### Required packages
```{r packages}
library(incidence2)
library(outbreaks)         # for example 1
library(covidregionaldata) # for example 2 and 3
library(clock)             # for example 3
library(zoo)               # for example 3
```

### Example 1 - Line list data

```{r example1}
# get the linelist data
dat <- ebola_sim_clean$linelist

# a quick visualising the daily onsets
dat |> incidence(date_of_onset) |> plot()

# calculate the epiweek incidence across hospitals and plot
dat |> 
  incidence(
    date_index = date_of_onset,
    groups = hospital,
    interval = "epiweek",
    na_as_group = FALSE
  ) |> 
  print() |> 
  facet_plot(nrow = 3, color = "white")
```

### Example 2 - Pre-aggregated data
```{r example2}
# get the data and filter out the nations
dat <- get_regional_data(country = "UK", verbose = FALSE)
nations <- c("England", "Northern Ireland", "Scotland", "Wales")
dat <- subset(dat, !region %in% nations)

# calculate iso-weekly new cases across regions and plot
dat |> 
  incidence(
    date_index = date,
    counts = cases_new,
    groups = region,
    interval = "isoweek",
    na_as_group = FALSE
  ) |> 
  print() |> 
  facet_plot(angle = 45)

```

### Example 3 - User supplied date grouping functions
```{r example3}
# month grouping based on clock functions
clock_month_inci <- 
  build_incidence(
    dat,
    date_index = date,
    groups = region,
    counts = ends_with("new"), # tidyselect :)
    FUN = function(x) calendar_narrow(as_year_month_day(x), precision = "month")
  )
clock_month_inci

# month grouping based on using zoo as.yearmon
zoo_month_inci <- 
  build_incidence(
    dat,
    date_index = date,
    groups = region,
    counts = ends_with("new"),
    FUN = as.yearmon
  )
zoo_month_inci

# comparing the above to the grates implementation
grates_month_inci <- 
  incidence(
    dat,
    date_index = date,
    groups = region,
    counts = ends_with("new"),
    interval = "month"
  )
grates_month_inci

# and check they are all "equal"
clock_month_inci$date_index <- as.Date(calendar_widen(clock_month_inci$date_index, "day"))
zoo_month_inci$date_index <- as.Date(zoo_month_inci$date_index)
grates_month_inci$date_index <- as.Date(grates_month_inci$date_index)

identical(
  as.data.frame(clock_month_inci),
  as.data.frame(zoo_month_inci)
)

identical(
  as.data.frame(grates_month_inci),
  as.data.frame(clock_month_inci)
)
```
