---
title: "Packages"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## incidence2
```{r, results="asis"}
cat("![](images/badge-maturing.svg)\n")
```
[incidence2](https://www.reconverse.org/incidence2/) Provides functions and
classes to compute, handle and visualise incidence from dated events for a
defined time interval. Dates can be provided in various standard formats. The
class 'incidence2' is used to store computed incidence and can be easily
manipulated, subsetted, and plotted.


## i2extras
```{r, results="asis"}
cat("![](images/badge-maturing.svg)\n")
```
[i2extras](https://www.reconverse.org/i2extras/) Provides additional
functionality to the [incidence2](https://www.reconverse.org/incidence2/)
package, including curve fitting, growth rate calculations and peak estimation.


## reportfactory
```{r, results="asis"}
cat("![](images/badge-maturing.svg)\n")
```
[reportfactory](https://www.reconverse.org/reportfactory/) facilitates workflows
for handling multiple .Rmd reports, compiling one or several reports in one go,
and storing outputs in well-organised, timestamped folders.


## trending
```{r, results="asis"}
cat("![](images/badge-maturing.svg)\n")
```
[trending](https://www.reconverse.org/trending/) provides a coherent interface
to specifying common models and facilitate the easy generation of confidence and
prediction intervals. Whilst it is useful in an interactive context, it’s main
focus is to provide an intuitive interface on which other packages can be
developed.


## trendeval
```{r, results="asis"}
cat("![](images/badge-experimental.svg)\n")
```
[trendeval](https://www.reconverse.org/trendeval/) provides an interface for
evaluating models fit with the trending package.












